Icicle: a single-pass query language

### 2015-12-06 2016icicle/section/02-Source.tex
ok

\TODO{Should we go through the evaluation rules here, or after typechecking?}
If we can write the evaluation rules to be ``agnostic'' of the type, it would be ideal to put them here.

@@

>>>>

\subsection{Evaluation}

@@

In figure~\ref{fig:source:eval} we describe the bigstep evaluation semantics of the language.
We use the shorthand $\ov{\Gamma}$ to mean a list of $\Gamma$, and this same overline syntax is used as a variable containing a list.
The judgment $\BigstepG{x}{\ov{v}}{v}$ uses a list $\ov{\Gamma}$, with one environment for each element of the input stream.
The environment $\Gamma$ is used for scalar variables that are not associated with any particular input element.
Next is the expression $x$ that is to be evaluated.
The last two are outputs $\ov{v}$ and $v$.
The output $\ov{v}$ is the intermediate result of the computation for each part of the input stream.
That is, each element of $\ov{v}$ is the result of evaluating $x$ on the input stream up to that element.
The final output, $v$, is the result at the end of the stream.
In most cases it is the same as the last element of $\ov{v}$, unless $\ov{v}$ is empty or for a variable lookup.

The reason for returning the result list $\ov{v}$ is to make implementing the @scan@ primitive easier.
This allows @scan@ itself to be evalauted effectively as a no-op, but does mean that each other rule has to implement its own @scan@ as well.

\TODO{It seems like we should be able to get rid of the single element return $v$, but then would it be strange with the return value being of length $max~1~|\ov{v}|$?}

When evaluating on an input such as our @stocks@ table, we would create an environment for each row, with open and close as the variables.
The rule (EvTop) simply uses the empty environment for the scalar environment, and ignores the stream output.

For variables, the rules (EvVar) and (EvVarInput) look up the variable in the contexts.
In the case of direct inputs as introduced by (EvTop), there is no corresponding entry in the scalar environment, so (EvVarInput) is used which returns an error as the scalar output.
In this way, if the input stream is used as an aggregate it will be an evaluation error, but if it is used as an element value and folded upon, the error value will be thrown away.
For normal variables, there should be a binding in all environments.

Let bindings are handled by rule (EvLet), which evaluates its definition, adds part of the stream result to each stream environments, and adds the scalar result to the scalar environment.

For the rule (EvFold), first the zero or initial value of the fold is computed, using only its scalar value.
The extra rules (ScanNil) and (ScanCons) are then used to repeatedly apply the fold computation to each stream input, starting from the initial value.
The list of fold values are added to the stream environments, and the last of the fold values, or the initial value if none, is added to the scalar environment.

In (EvFilter) the predicate is evaluated to a list of results. 
The stream environments are filtered according to this predicate list using $\mit{pack}$, and the rest of the expression is computed.
However, the result list of the rest of the expression is only the scan of the filtered stream inputs.
We use $\mit{extend}$ to add back parts of the stream that have been filtered out, but must first find the initial value of the scan to start from.

The primitive operator rules (EvAdd) and (EvGt) apply their operator to each element of the inputs.
The primitive (EvNat) returns a list of constant numbers, one for each stream element.
Finally, (EvScan) is a no-op since the hard work of computing the @scan@s is actually implemented in the rest of the rules.
